Red Oak: Proactive Policy Generation

### 2015-05-09 draft.tex
..

Figure~\ref{fig:firewall-eg} displays a policy for a simple stateful firewall using our API. For simplicity, we assume that IP ranges are between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
In the example, source IPs from 0 to 10 are considered to be internal to the network and everything else, external.

@@

 For example, after having seen packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:

@@

type forwarding_decision =
	| Deliver
	| Drop
	| ForwardTo of int
	| Multicast of forwarding_decision * forwarding_decision

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
   % 2) Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.

>>>>

A policy for a simple stateful firewall using our API is included in Figure~\ref{fig:firewallcode}. For simplicity, we assume that IP ranges are between 0 and 1000, however it is easy to generalize this to actual IP prefixes.
In the example, source IPs from 0 to 10 are considered to be internal to the network and everything else, external.

@@

 For example, after having seen a packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the first rules would be matched, evaluating the decision tree would add tuple $(10,150)$ to relation \lstinline|conn|, the packet would be forwarded according to the forwarding decision \lstinline|Deliver|, and the concrete rules would updated as follow:

@@

type forwarding_decision =
	| Deliver
	| Drop
	| ForwardTo of int
	| Multicast of forwarding_decision
                     l* forwarding_decision

@@

\subsection*{Optimizations}
   % 1) expand tree before add/remove nodes with a match on that relation ( based on add(p); add(p) == add(p) so can shortcircuit if p is already in the relation)
The rules generated by the presented technique ensures that the only packets being forwarded to the controller are those reaching an add or a remove node and thus potentially modify the state of the controller. However, these packets do not always modify the state of the controller. For example, in the firewall example (see Fig~\ref{fig:firewallcode}), after seeing a first packet with an \lstinline|ipsrc| of 10 and \lstinline|ipdst| of 150, any subsequent packet with the same \lstinline|ipsrc| and \lstinline|ipdst| could be delivered without being first forwarded to the controller, as adding the same tuple to a relation is idempotent. By using both local and global transformations on the decision tree, we can get closer to an optimal amount of packet forwarded to the controller, where only packets modifying relations are seen by the controller.

We implemented in Red Oak, under \lstinline|o1|, a local transformation adding an \lstinline|in_rel| node before \lstinline|add| or \lstinline|remove| nodes. In the case of \lstinline|add|, the \lstinline|add| node is omited in the true branch of \lstinline|in_rel|, such that packets whose fields are already in the relation will not be seen by the controller due to this \lstinline|add| node. Similarly, we omit the \lstinline|remove| node in the false branch of \lstinline|in_rel|. In the firewall example (see Fig.~\ref{fig:firewallcode}), this would result the following decision tree:

\begin{figure}[ht]
  TODO
\caption{Optimized Decision Tree for Stateful Firewall}     
\label{fig:decisiontreeopt}  
  \end{figure}

From which the following symbolic rule would be generate:
\begin{lstlisting}
ipsrc in [0, 100], (ipdst, ipsrc) in conn
                  --> Deliver
ipsrc in [0, 100] --> Sent to Controller
ipsrc in [101, 1000], (ipsrc,ipdst) in conn
                  --> Deliver
ipsrc in [101, 1000], (ipsrc,ipdst) not in conn
                  --> Drop
\end{lstlisting}

With empty relation, the concrete rules would be the same as with the unoptimized decision tree. However, after seeing, for example, packet with \lstinline|ipsrc| 10 and \lstinline|ipdst| 150, the rules would be
\begin{lstlisting}
ipsrc = 10, ipdst = 150 ---> Deliver  
ipsrc in [0, 100] ---> Sent to Controller  
ipsrc = 150, ipdst = 10 ---> Deliver  
ipsrc in [101, 1000] ---> Drop  
\end{lstlisting}

   % 2) Prune dead subtree on empty ranges and add/remove which do not affect any inhabitant-based branching.
